---
title: "Juvenile Salmon Mainstem Monitoring Program"
params:
  runs:  #just as a placeholder. the render() function will overwrite this params on YAML
  - Fall
  - Winter 
  year:
  - 2011
  - 2012
output:
  pdf_document: default
geometry: margin=0.5in    
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
#- \usepackage{xcolor} 
- \usepackage{subcaption}
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = '', fig.width = 8, fig.height = 6, 
  warning = FALSE, fig.align = "center", message = FALSE)
```
## Introduccion
NOTE: THIS rmd file is run by the 'report.R' function or it can be ran directly from the knit menu here.

A statistical analysis was conducted to determine which environmental covariates
were related to the efficiency of rotary screw traps operated at the Red Bluff Diversion Dam
trap site on the Sacramento River Mainstem. That analysis suggests that LunarPhase, Mean-
FLRelease andWaterTemperature were not significantly related to trap efficiency. In contrast,
the analysis suggests that PercQ, Turbidity, RiverDepth andWeather were significantly related
to trap efficiency.


```{r, echo=FALSE}
# Packages we will use
#if(!require("pacman")){
 # install.packages("pacman")
##}
#pacman::p_load(tidyverse, table1, compareGroups, kableExtra,readxl,janitor,gt)
library(tidyverse)
library(kableExtra)
library(readxl)
library(table1)
library(scales)
library(janitor)
library(gt)
library(patchwork)
library(parseLatex)

salmon <- read_excel('RBsalmon.xlsx')
salmon <- na.omit(salmon)
salmon$SampleDate <- as.Date(salmon$SampleDate,"%m/%d/%Y")
salmon <- dplyr::filter(salmon,!race == "n/p") 
salmon <- dplyr::filter(salmon,ForkLength > 0) 
salmon$year <- format(salmon$SampleDate,"%Y") #year
salmon$week <- format(salmon$SampleDate,"%V") #week starts with 01 on jan 1..also "%W" and "%U" can be used
salmon$month <- format(salmon$SampleDate,"%B") #month name..abreviated "%b" or number "%m"
```

```{r, echo=FALSE}

 salmon <- salmon[,c(1,2,5,6,9,11,13,14:16)]

#salmon <- salmon %>% dplyr::filter(year >2009 & year < 2012); unique(salmon$year)

filtered_data <- salmon %>%
  dplyr::filter(year %in% params$year, race2 %in% params$runs) #%>% #Must use %in% operator, otherwise, == will give the wrong Count per run 
 #arrange(month, SampleDate)

#Get the total count for all the runs in current year
#all_runs <- salmon %>% dplyr::filter(year %in% params$year) %>% 
 #group_by(year, race2) %>%
 # summarise(total = sum(Count))

#Subsetting works best than filtering with dplyr because kable was only giving me the first column
all_runs <- subset(salmon, year %in% params$year)
all_runs <- all_runs %>% 
  group_by(year, race2) %>%
  summarise(total = sum(Count))

#make it a data.frame or table won't print properly
all_runs2 <- data.frame(all_runs)

total_run <- sum(all_runs$total)
total_run <- format(total_run, big.mark = ",", scientific = FALSE)

#Put commas on total..NOTE: I didnt put commas an 'all_runs' above because it would throw an error when summing 'total_run'
all_runs2 <- all_runs2 %>% mutate(total = format(total, big.mark = ",", scientific = FALSE))
```

```{r, echo =FALSE, results='asis'}
#Now get the total count for the current run and year
#a <- salmon[,c(1,2,5,6,9,11,13,14:16)]
this_run <- salmon %>% dplyr::filter(year %in% params$year & race2 %in% params$runs) %>% 
 group_by(race2, year) %>%
  summarise(total = sum(Count))

total_catch <- sum(this_run$total)
total_catch <- format(total_catch, big.mark = ",", scientific = FALSE)

#Put commas on total..NOTE: I didnt put commas an 'this_run' above because it would throw an error when summing 'total_catch'
this_run <- this_run %>% mutate(total = format(total, big.mark = ",", scientific = FALSE))
runs <- unique(all_runs$race2)

#Caption for summarized side by side tables
caption_text1 <- paste("This table represents", paste(params$runs, collapse = ", "), "juvenile salmon run (s) for", paste(params$year, collapse = ", "), "brood year (s)")

caption_text2 <- paste("This table represents the total", paste(runs, collapse = ", "), "juvenile salmon run (s) for", paste(params$year, collapse = ", "), "brood year (s)")

t1 <- kbl(this_run, booktabs = TRUE, format = 'latex', linesep = "") %>% 
  kable_styling(latex_options = c("striped", "Hold_position"), stripe_color = "cyan!10")

t2 <- kbl(all_runs2, booktabs = TRUE, format = 'latex', linesep= "") %>% 
  kable_styling(latex_options = c("striped"), stripe_color = "pink!25")

t1 = substr(t1, start = 25, stop = nchar(t1) - 11) ##USE this to avoid crashes when using kable_styling from kableExtra (see chunk above).
t2 = substr(t2, start = 25, stop = nchar(t2) - 11) ##USE this to avoid crashes when using kable_styling from kableExtra (see chunk above).

cat(sprintf("
\\begin{table}[H]
    \\caption{Juvenile Salmon Outmigrating through the Mainstem Diversion Dam}
    \\centering
    \\begin{subtable}[t]{0.48\\textwidth}
        \\centering
        \\caption{%s}
        %s   
    \\end{subtable}%%
    \\hfill
    \\begin{subtable}[t]{0.48\\textwidth}
        \\centering
        \\caption{%s}
        %s
    \\end{subtable}
\\end{table}
", caption_text1, t1, caption_text2, t2))
```

```{r, echo=FALSE}
#run <- salmon %>% filter(year %in% params$year & race2 %in% params$runs)
#head(run, 20)
```

```{r, echo=FALSE}
library(kableExtra)
kbl(head(filtered_data, 15), booktabs = T, linesep = "", caption = paste("This table represents", paste(params$runs, collapse = ", "), "and", paste(params$year, collapse = ", "), "brood year (s)")) %>%
  #linesep removes the white space every 5 rows that prints in kableextra
kable_styling(latex_options = "HOLD_position") %>% #HOLD_position option doesn't work when striped is included in kable_styling
  #NOTE: TAKE A LOOK AT THE xcolor PACKAGE TO SEE AVAILABLE COLORS FOR kableExtra TABLES AND BE ABLE TO USE SOMETHING LIKE "violet!5"
kable_styling(latex_options = "striped", stripe_color = "violet!5")  #I had to include another kable_styling because striped wouldn't work when HOLD_position was in place
```


The total catch of `r params$runs` is `r total_catch` for brood year(s) `r params$year` and a grand total of `r total_run` for all the runs `r runs` combined. 

```{r,echo=FALSE, cache=FALSE}
#De-aggregate to convert Count into single rows
d_ag <- filtered_data[rep(1:nrow(filtered_data),filtered_data$Count),]
d_ag$Count2 <- rep(1,nrow(d_ag))

run_sum <- d_ag %>%  
   group_by(race2, year) %>%
    summarise(total = sum(Count2))

#Extract fish columns only to be used below dynamically
fish_cols <- unique(d_ag$race2)
```

```{r, echo=FALSE}
 #Make a pivot_wider so the salmon Runs will be headers for gt
salmon_gt <- d_ag %>% 
  group_by(race2, year) %>% 
  mutate(obs = row_number()) %>% 
  pivot_wider(names_from = "race2", values_from = "Count2", values_fill = 0) %>% 
  arrange(year, obs) %>% 
  select(-obs) 
#dim(salmon_gt)
#salmon_gt
```

## Summary Tables

```{r echo=FALSE, cache=FALSE}
library(table1)
comma_categorical <- function(x) {
  c("", sapply(stats.default(x), function(y) with(y,
    sprintf("%s (%s%%)", 
            prettyNum(FREQ, big.mark = ","),
            format(round(PCT, 1), nsmall = 1)))))
}

x <- table1(~ race2 + StationCode + year + ForkLength, data = d_ag, render.categorical = comma_categorical, droplevels = F, overall = "Overall" ,caption="Summary Table")

t1kable(x) |>
kable_styling(latex_options = "HOLD_position") %>% 
kable_styling(latex_options = "striped")

#IMPORTANT: SampleDate MUST be converted to 'character' or table1 will crash if format as.Date
#SO, just for this table convert SampleDate to character but convert back to as.Date if needed
d_ag$SampleDate <- as.character(d_ag$SampleDate)

xx <- table1(~ StationCode + ForkLength | race2, data = d_ag, render.categorical = comma_categorical, overall = "Overall" ,caption="Juvenile salmon migrating out to the ocean")

t1kable(xx) |>
kable_styling(latex_options = "HOLD_position") %>% 
kable_styling(latex_options = "striped")

```

```{r, echo=FALSE}
salmon_gt <- salmon_gt |> dplyr::mutate(week = strftime(SampleDate, format = "%V"))  #week starts with 01 on jan 1..also "%W" and "%U" can be used
#salmon_d_ag$week <- with(salmon_d_ag,ifelse(month=='December' & day==31,"52",week))
#salmon_gt

#Break salmon_gt into subsets of weeks to be able to print headers on every page of longtable
#gt_52_36 <- salmon_gt %>% filter(week > 35)
#gt_52_36 |> 
 

###Plot only a few months for simplicity
salmon_gt <- salmon_gt %>% dplyr::filter(month > 'May')
  #source('repeat_header_gt.R')

salmon_gt |>
  mutate(SampleDate = as.character(SampleDate)) |>
  group_by(year, month) |>   #MANIPULATE group TO CHANGE THE TABLES SUMMARIZATION
  summarise(across(all_of(fish_cols),sum), across(c(Turbidity,WaterTemperature), mean)) %>%  #Notice the double across call below...GREAT!!!
  rowwise() %>% mutate(Total = sum(across(all_of(fish_cols)))) |> 
  #summarise(across(c(Winter, Fall, Spring), sum), across(c(Turbidity,WaterTemperature), mean)) |> 
  
arrange(desc(year),desc(month)) |>  
  gt() |> tab_caption(caption = "Nice longtable with headers on every page") %>% #Use tab_caption to autonumber gt tables
fmt_number(decimals = 0, sep_mark = ",") |>
  
summary_rows( # sum WINTER, FALL, LATEFALL and SPRING ONLY
    columns = c(fish_cols, Total),
    #columns = c(Winter, Spring, Fall),
    fns =  list(label = "Total", fn = "sum"), side = "bottom", fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  
  summary_rows( # GET THE mean OF TURBIDITY AND WATERTEMPERATURE ONLY
    columns = c(Turbidity,WaterTemperature),
    fns =  list(label="avg",fn = "mean"),
   fmt = ~ fmt_number(., decimals = 1, use_seps = FALSE)) |> #Format the 'avg' row
  #round Turbidity and WaterTemperature columns to one decimal
  fmt_number(columns = c(Turbidity,WaterTemperature),
             decimals = 1) |>

 # grand_summary_rows(columns = -c(SampleDate,Turbidity,WaterTemperature),
    grand_summary_rows(columns = c(fish_cols, Total),
    #grand_summary_rows(columns = c(Winter, Spring, Fall),
  fns = list(label = "Grand Total", fn = "sum"), ,fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  tab_header(title = "Data summarized by year and month",
    subtitle = "Total number of fish by year and month") %>% 
  tab_options(latex.use_longtable = TRUE) %>%
            BD::repeat_header_gt() #This option prints headers on every single page for longtable(NOT WORKING RIGHT NOW)
```

```{r, echo=FALSE}
salmon_gt <- salmon_gt |> dplyr::mutate(week = strftime(SampleDate, format = "%V"))  #week starts with 01 on jan 1..also "%W" and "%U" can be used

###Plot only a few months for simplicity
#salmon_gt <- salmon_gt %>% dplyr::filter(month > 'May')
  #source('repeat_header_gt.R')

salmon_gt |>
  mutate(SampleDate = as.character(SampleDate)) |>
  group_by(year, month, week) |>   #MANIPULATE group TO CHANGE THE TABLES SUMMARIZATION
  summarise(across(all_of(fish_cols),sum), across(c(Turbidity,WaterTemperature), mean)) %>%  #Notice the double across call below...GREAT!!!
  rowwise() %>% mutate(Total = sum(across(all_of(fish_cols)))) |> 
  #summarise(across(c(Winter, Fall, Spring), sum), across(c(Turbidity,WaterTemperature), mean)) |> 
  
arrange(desc(year),desc(month), desc(week)) |>  
  gt() |> tab_caption(caption = "Long Table of fish caught by year, month and week") %>% #Use tab_caption to autonumber gt tables
fmt_number(decimals = 0, sep_mark = ",") |>
  
summary_rows( # sum WINTER, FALL, LATEFALL and SPRING ONLY
    columns = c(fish_cols, Total),
    #columns = c(Winter, Spring, Fall),
    fns =  list(label = "Total", fn = "sum"), side = "bottom", fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  
  summary_rows( # GET THE mean OF TURBIDITY AND WATERTEMPERATURE ONLY
    columns = c(Turbidity,WaterTemperature),
    fns =  list(label="avg",fn = "mean"),
   fmt = ~ fmt_number(., decimals = 1, use_seps = FALSE)) |> #Format the 'avg' row
  #round Turbidity and WaterTemperature columns to one decimal
  fmt_number(columns = c(Turbidity,WaterTemperature),
             decimals = 1) |>

 # grand_summary_rows(columns = -c(SampleDate,Turbidity,WaterTemperature),
    grand_summary_rows(columns = c(fish_cols, Total),
    #grand_summary_rows(columns = c(Winter, Spring, Fall),
  fns = list(label = "Grand Total", fn = "sum"), ,fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  tab_header(title = "Data summarized by year, month, and week",
    subtitle = "Total number of fish by year, month, and week") %>% 
  tab_options(latex.use_longtable = TRUE) %>%
            BD::repeat_header_gt() #This option prints headers on every single page for longtable(NOT WORKING RIGHT NOW)
```

\newpage
```{r, echo=FALSE}
salmon_gt <- salmon_gt |> dplyr::mutate(week = strftime(SampleDate, format = "%V"))  #week starts with 01 on jan 1..also "%W" and "%U" can be used

###Plot only a few months for simplicity
#salmon_gt <- salmon_gt %>% dplyr::filter(month > 'May')
  #source('repeat_header_gt.R')

salmon_gt |>
  mutate(SampleDate = as.character(SampleDate)) |>
  group_by(year, week) |>   #MANIPULATE group TO CHANGE THE TABLES SUMMARIZATION
  summarise(across(all_of(fish_cols),sum), across(c(Turbidity,WaterTemperature), mean)) %>%  #Notice the double across call below...GREAT!!!
  rowwise() %>% mutate(Total = sum(across(all_of(fish_cols)))) |> 
  #summarise(across(c(Winter, Fall, Spring), sum), across(c(Turbidity,WaterTemperature), mean)) |> 
  
arrange(desc(year), desc(week)) |>  
  gt() |> tab_caption(caption = "Long Table of yearly catch by week") %>% #Use tab_caption to autonumber gt tables
fmt_number(decimals = 0, sep_mark = ",") |>
  
summary_rows( # sum WINTER, FALL, LATEFALL and SPRING ONLY
    columns = c(fish_cols, Total),
    #columns = c(Winter, Spring, Fall),
    fns =  list(label = "Total", fn = "sum"), side = "bottom", fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  
  summary_rows( # GET THE mean OF TURBIDITY AND WATERTEMPERATURE ONLY
    columns = c(Turbidity,WaterTemperature),
    fns =  list(label="avg",fn = "mean"),
   fmt = ~ fmt_number(., decimals = 1, use_seps = FALSE)) |> #Format the 'avg' row
  #round Turbidity and WaterTemperature columns to one decimal
  fmt_number(columns = c(Turbidity,WaterTemperature),
             decimals = 1) |>

 # grand_summary_rows(columns = -c(SampleDate,Turbidity,WaterTemperature),
    grand_summary_rows(columns = c(fish_cols, Total),
    #grand_summary_rows(columns = c(Winter, Spring, Fall),
  fns = list(label = "Grand Total", fn = "sum"), ,fmt = ~ fmt_number(.,decimals = 0, sep_mark = ",")) |>
  tab_header(title = "Data summarized by year and week",
    subtitle = "Total number of fish by year and week") %>% 
  tab_options(latex.use_longtable = TRUE) %>%
            BD::repeat_header_gt() 
```

